---
title: "CRUK CI Bitesize Introduction to Single Cell RNASeq Analysis"
subtitle: 'Dataset Integration and Batch Correction - clinic session'
author: "Abbi Edwards"
output:
  html_document:
    df_print: paged
    toc: yes
    number_sections: true
    code_folding: show
    css: ../css/boxes.css
---

# Data integration - PBMMC and ETV6-RUNX samples

```{r library_PBMMC_ETV6-RUNX1, warning=FALSE, message=FALSE}
library(scater)
library(scran)
library(batchelor)
library(bluster)
library(tidyverse)
library(pheatmap)
library(BiocSingular)
```

## Load the data

We are loading a single `sce` object containing the 7 Caron samples. The data has been processed as detailed in previous weeks and downsampled to 500 cells per sample for demonstration purposes.

```{r, loadSamples}
all.sce <- readRDS("../Robjects/DataIntegration_all_sce_dimred.Rds")
all.sce
```

## View the uncorrected data

Before we actually perform any correction, it is worth examining whether there is any batch effect in this dataset.The PCA, TSNE and UMAP have already been run but we can do some quick clustering just so we can see things a bit better and this will be needed for the diagnostics later.

```{r}
uncorrected.snn.gr <- buildSNNGraph(all.sce, use.dimred="PCA")
uncorrected.clusters <- igraph::cluster_walktrap(uncorrected.snn.gr)$membership
colLabels(all.sce) <- factor(uncorrected.clusters)
all.sce$Cluster <- factor(uncorrected.clusters)
uncorrected.tab <- table(Cluster=uncorrected.clusters, Batch=all.sce$SampleName)
uncorrected.tab
```


```{r}
plotTSNE(all.sce, colour_by="SampleName")
```

Reminder: Of course, the other explanation for batch-specific clusters is that there are cell types that are unique to each batch. 

## Correct the data - quickCorrect wrapper

```{r}
quick.corrected.all <- quickCorrect(all.sce, batch = all.sce$SampleName)

quick.sce.all <- quick.corrected.all$corrected

quick.sce.all
```

## Visualise the corrected data

```{r}
set.seed(0010101011)
quick.sce.all <- runTSNE(quick.sce.all, dimred="corrected")
quick.sce.all$SampleName <- factor(all.sce$SampleName)
plotTSNE(quick.sce.all, colour_by="SampleName")
```


## Retrieve the metadata

```{r}
meta <- colData(all.sce) %>%
  as.data.frame() %>%
  DataFrame()

colData(quick.sce.all) <- meta
  
```

## Correction Diagnostics

### Mixing Between Batches

Get the cluster information 

```{r}
all.snn.gr <- buildSNNGraph(quick.sce.all, use.dimred="corrected", k=20)
clusters.all <- igraph::cluster_walktrap(all.snn.gr)$membership
colLabels(quick.sce.all) <- factor(clusters.all)
quick.sce.all$Cluster <- factor(clusters.all)

clusterTab.all <- data.frame(clusters=clusters.all, batch=quick.sce.all$SampleName, source=quick.sce.all$SampleGroup)

all.tab <- table(Cluster=clusters.all, Batch=quick.sce.all$SampleName)
all.tab
```

```{r}
plotTSNE(quick.sce.all, colour_by="Cluster") + 
  facet_wrap(~colData(quick.sce.all)$SampleName)
```

```{r}
ClusterInfo.all <- data.frame(Cluster=clusters.all, Batch=quick.sce.all$SampleName, source=quick.sce.all$SampleGroup) %>%
  group_by(Cluster,Batch) %>%
  summarise(cells = n())

ggplot(data=ClusterInfo.all, aes(x=Cluster,y=cells, fill=Batch)) +
	theme(axis.text.x=element_text(angle = 90, hjust = 0)) +
	geom_col() +
	theme(legend.text = element_text(size = 7))

```


```{r}
var <- clusterAbundanceVar(all.tab)

# Also printing the percentage of cells in each cluster in each batch:
percent <- t(t(all.tab)/colSums(all.tab)) * 100 
df <- DataFrame(Batch=unclass(percent), var=var)
df[order(df$var, decreasing=TRUE),]

```

### Preserving Biological Heterogeneity

Comparing the clustering pre and post correction

```{r}

# PBMMC 1b

only_PBMMC1b <- all.sce[,all.sce$SampleName == "PBMMC_1b"]

tab.sample.PBMMC1b <- nestedClusters(ref=paste("before", colLabels(only_PBMMC1b)),
    alt=paste("after", clusters.all[quick.sce.all$SampleName=="PBMMC_1b"]))
tab.sample.PBMMC1b$alt.mapping
```

```{r}
# ETV6-RUNX1_1

only_ETV6RUNX1_1 <- all.sce[,all.sce$SampleName == "ETV6-RUNX1_1"]

tab.sample.ETV6RUNX1_1 <- nestedClusters(ref=paste("before", colLabels(only_ETV6RUNX1_1)),
    alt=paste("after", clusters.all[quick.sce.all$SampleName=="ETV6-RUNX1_1"]))
tab.sample.ETV6RUNX1_1$alt.mapping
```

We can visualize this mapping for the samples. 

```{r}
# For the first sample:
heat.sample.1 <- pheatmap(tab.sample.PBMMC1b$proportions, cluster_row=FALSE, cluster_col=FALSE,
                   main="Sample 1 comparison", silent=TRUE)

# For the second sample:
heat.sample.2 <- pheatmap(tab.sample.ETV6RUNX1_1$proportions, cluster_row=FALSE, cluster_col=FALSE,
                   main="Sample 2 comparison", silent=TRUE)

gridExtra::grid.arrange(heat.sample.1[[4]], heat.sample.2[[4]])
```

## Adjusted Rand index 

```{r}
ri.sample.1 <- pairwiseRand(clusters.all[quick.sce.all$SampleName=="PBMMC_1b"], colLabels(only_PBMMC1b), mode="index")
ri.sample.1
```

```{r}
ri.sample.2 <- pairwiseRand(clusters.all[quick.sce.all$SampleName=="ETV6-RUNX1_1"], colLabels(only_ETV6RUNX1_1), mode="index")
ri.sample.2
```

ARI on per-cluster level

```{r}
# For the first batch.
tab1 <- pairwiseRand(uncorrected.clusters[all.sce$SampleName=="PBMMC_1b"], clusters.all[quick.sce.all$SampleName=="PBMMC_1b"])
heat.1 <- pheatmap(tab1, cluster_row=FALSE, cluster_col=FALSE,
    col=rev(viridis::magma(100)), main="PBMMC_1b probabilities", silent=TRUE)

# For the second batch.
tab2 <- pairwiseRand(uncorrected.clusters[all.sce$SampleName=="ETV6-RUNX1_1"], clusters.all[quick.sce.all$SampleName=="ETV6-RUNX1_1"])
heat.2 <- pheatmap(tab2, cluster_row=FALSE, cluster_col=FALSE,
    col=rev(viridis::magma(100)), main="ETV6-RUNX1_1 probabilities", silent=TRUE)

gridExtra::grid.arrange(heat.1[[4]], heat.2[[4]])

```

### MNN specific test

lost variance

```{r}
metadata(quick.sce.all)$merge.info$lost.var
```


## Session information

```{r sessionInfo_dsi_PBMMC_ETV6-RUNX1}
sessionInfo()
```